% Place acknowledgments below.
The completion of graduate school and this dissertation could not have been possible without the support from many people that I would like to thank now.  
First, I would like to thank Prof. Gary Messing for guiding this research and helping me develop as a scientist. I learned a lot from Prof. Messing and I will always be indebted to him for his time, patience, and experience. 
I would like to thank my dissertation committee, Prof. James Adair, Prof. Douglas Wolfe, Prof. John Hellmann and Dr. Charles Compson for their time, constructive criticism and advice. 
I would like to thank Dr. Charles Compson and Dr. Joe Atria from Almatis, Inc. for funding this research and for many discussions that contributed substantially to the quality of this research. 
The staff of the Materials Characterization Laboratory was always supportive and helpful and I enjoyed working with them. Specifically, I would like to thank Joshua Maier and Haiying Wang for preparing my samples for TEM, and Dr. Jennifer Gray for spending many hours at the TEM with me.
I would like to thank the past and present members of the Messing Research Group, (Dr. Elizabeth Kupp, Jesse Matson, Alexandra Wadja, Elisabeth Roch, Ryan Sommerhuber, Dr. Nuerxida Pulati, Marie Arnoult, Yi Yuan, Dr. Yunfei Chang, Beecher Watson, Rebecca Walton, Michael Brova, Anna Hofer, Kuo-Hao Lee, Jie Wu, Howard Payne, Dr. Raul Bermejo) for a great time in the lab and in the office, and many great discussions. It was an honor to get to work with so many great people.
I want to thank the staff, faculty, and students of the Penn State Materials Science and Engineering Department for always being helpful and for creating a positive and great work environment.
I want to thank all the friends I made at Penn State over the past few years, within and outside the Department of Materials Science (Christopher Rotella, Lyndsey Denis, Daniel Schuman, Nathan Smith, Tim Lichtenstein, Ethan Kahn, Nathan Smith, Tim Pillsbury, and many more). I am proud that I can call them my friends.
I would like to thank my significant other, Cassie Marker for her support. She also contributed significantly to this dissertation by performing the first-principles DFT calculations, which I could not have done without her.
Finally, I want to thank my family. My parents, grandparents, uncle, and brother always supported and encouraged me, even when I decided to move half way across the globe for grad school. I am very grateful for their constant support.
